Making multiple-locus association mapping on a genome-wide scale routine

### 2016-01-29 AM+draft.tex
2901 Finalized version of changed paragraph in Indroduction

Over the past decade, there have also been impressive advancements in genome wide association mapping methods, the statistical techniques for measuring the strength of association between a marker locus and trait. Traditionally, association data were analysed with simple contingency tests or Cochran-armitage  trend tests. Today, far more statistically sophisticated methods are available. One such method is the  QK approach, a linear mixed model framework for genome wide association mapping (refs).  Several highly efficient 
computer programs for performing QK-based association mapping are available. Programs such as EMMA, fastLMM, GEMMA, and 
XXX(ref) make feasible the analysis of data from hundreds of thousands of snps, collected on thousands of subjects. However, not all 
areas of genome-wide association mapping have advanced. It still remains standard practice to analyse genome-wide association data 
by measuring the strength of association between a marker locus and trait for each locus separately.

Over the past decade, there have also been impressive advancements in genome wide association mapping methods. Association mapping methods are statistical techniques for measuring the strength of association between a marker locus and trait. It has long been recognised that spurious marker-trait associations can occur due to population structure. This problem has now been solved (refs). A linear mixed model framework has also been shown to be well suited to analysing data from gwas. Linear mixed models can accommodate multiple sources of variation simultaneously, including population structure and familial relatedness (refs). Fitting lmm's to gwas data though was computationally challenging. This limitation has also been solved via spectral decomposition (ref) and approximation (ref) leading to several very fast lmm implementations (refs). Despite these impressive advancements, a core feature of mainstream  association mapping has remained unchanged. The strength of association between a maker locus and trait is assessed via a single locus model.  That is, the trait and marker data from gwas are analysed for each marker locus separately.

>>>>

Over the past decade, there have also been impressive advancements in genome wide association mapping methods, the statistical techniques for measuring the strength of association between a marker locus and trait. Traditionally, association data were analysed with simple contingency tests or Cochran-armitage  trend tests. Today, far more statistically sophisticated methods are available. One such method is the  QK approach, a linear mixed model framework for genome wide association mapping (refs).  Several highly efficient 
computer programs for performing QK-based association mapping are available. Programs such as EMMA, fastLMM, GEMMA, and 
XXX(ref) make feasible the analysis of data from hundreds of thousands of snps, collected on thousands of subjects. However, not all 
areas of genome-wide association mapping have advanced. It still remains standard practice to analyse genome-wide association data with a single locus model. That is,  the strength of association between a marker locus and trait is measured 
for each locus separately.
